---
title: "PrepR Summary Report"
format: PrettyPDF-pdf
fig.height: 4.5
fig.width: 6.5
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
require(tidyverse)
require(magrittr)
require(terra)
require(sf)
require(pals)
require(ochRe)
require(showtext)
require(ggtext)
require(gghighlight)
require(ochRe)
require(ggspatial)
require(gt)
require(tidyterra)
require(ggpmisc)
require(viridis)

```

```{r graphics-setup, include = FALSE}

pal.parl <- colorRampPalette(ochRe::ochre_palettes[["parliament"]])
pal.nam <- colorRampPalette(ochRe::ochre_palettes[["namatjira_qual"]])
font_add_google('Anton', 'anton')
showtext_auto()

```

This report provides several metrics of interest over the geographic extent of the AOI, including the distribution of forest age, dominant species, common disturbances, topographic variables, and others. It also summarizes key outputs from the prepR workflow and briefly describes their purpose in the modelling process. More details on the data, methodology, and results from the prepR workflow can be found in the \[PEM Manual\].

```{r aoi-map}
# 
# aoi2 <- st_read("C:/Users/clarmour/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/aoi.gpkg")
# vri2 <- st_read("C:/Users/clarmour/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/vri.gpkg")

aoi <- sf::st_read("C:/Users/clarmour/OneDrive - Government of BC/UBC/phd_data/vectors_AOI/BDY_M_AOI.shp")
vri <- sf::st_read("C:/Users/clarmour/OneDrive - Government of BC/UBC/phd_data/VRI/VEG_R1_PLY_BDY_R.shp")

# vri_int <- st_join(vri, aoi, join = st_intersects)
# 
# bec <- vri_int %>%
#   dplyr::mutate(BEC_LABEL = str_c(BEC_ZONE_C, BEC_SUBZON),
#                 BEC_LABEL = case_when(
#                   !is.na(BEC_VARIAN) ~ str_c(BEC_LABEL, BEC_VARIAN),
#                   TRUE ~ BEC_LABEL))

```

```{r aoi-inset, echo = FALSE}

#| fig-cap: "Figure 1: A map of where the [AOI_name] AOI is located in British Columbia."
#| fig-width: 6
#| fig-height: 4

aoi <- sf::st_read("C:/Users/ccarmour.stu/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/aoi.gpkg", quiet = TRUE)

bc_outline <- sf::st_read("C:/Users/ccarmour.stu/Documents/GitHub/PEMtestr/summary_report/bc_outline.gpkg", quiet = TRUE)

ggplot2::ggplot() +
  tidyterra::geom_spatvector(data = terra::vect(bc_outline), fill = "floralwhite") +
  tidyterra::geom_spatvector(data = terra::vect(aoi), color = "firebrick", fill = "floralwhite", linewidth = 1) +
  ggspatial::annotation_north_arrow(
    location = "tr", # Position: bottom-right
    which_north = "true", # True North
    pad_x = unit(0.5, "cm"), # Padding from x-axis
    pad_y = unit(0.5, "cm"), # Padding from y-axis
    style = ggspatial::north_arrow_fancy_orienteering() # Style of the arrow
  )

```

```{r bec-map, echo = FALSE}

#| fig-cap: "Figure 2: A map of BEC subzones, variants, and phases across [AOI_name]."
#| fig-width: 6
#| fig-height: 4

bec <- st_read("C:/Users/ccarmour.stu/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/bec.gpkg", quiet = TRUE) %>%
  dplyr::mutate(BGC_LABEL = str_replace_all(BGC_LABEL, " ", ""))

ggplot(data = terra::vect(bec), mapping = aes(fill = BGC_LABEL)) +
  tidyterra::geom_spatvector(linewidth = 0) +
  labs(fill = "BEC Subzone/Variant") +
    ggspatial::annotation_north_arrow(
    location = "tr", # Position: bottom-right
    which_north = "true", # True North
    pad_x = unit(0.5, "cm"), # Padding from x-axis
    pad_y = unit(0.5, "cm"), # Padding from y-axis
    style = ggspatial::north_arrow_fancy_orienteering() # Style of the arrow
  )

```

```{r bec-chart, echo = FALSE}

#| fig-cap: "Table 1: A summary of BEC subzones and variants by land area."
#| fig-width: 6
#| fig-height: 4

bec.df <- bec %>%
  dplyr::mutate(AREA.METRES = sf::st_area(.)) %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(BGC_LABEL) %>%
  dplyr::summarize(AREA.BY.GRP = round(as.numeric(sum(AREA.METRES, na.rm = TRUE))/10000000, 2))

bec.table <- bec.df %>%
  gt::gt() %>%
  tab_header(title = "BEC Subzone/Variant by Land Area") %>%
  cols_label(
    BGC_LABEL = "Subzone/Variant",
    AREA.BY.GRP = "Area (1000s of ha)"
  )

```

```{r forest-age, echo = FALSE}

#| fig-cap: "Figure 3: Distribution of forest age across [AOI_name], rounded to nearest decade. Polygons with an age of 250 years or greater were clamped to 250 years for visualization purposes."
#| fig-width: 6
#| fig-height: 4

vri <- sf::st_read("C:/Users/ccarmour.stu/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/vri_detail.gpkg", quiet = TRUE) %>%
  sf::st_join(bec, join = st_intersects, quiet = TRUE, left = FALSE) %>%
  dplyr::mutate(BEC_LABEL = str_c(BEC_ZONE_CODE, BEC_SUBZONE),
                BEC_LABEL = case_when(
                  !is.na(BEC_VARIANT) ~ str_c(BEC_LABEL, BEC_VARIANT),
                  TRUE ~ BEC_LABEL),
                PROJ_AGE_1 = case_when(
    PROJ_AGE_1 == 0 ~ NA_integer_,
    PROJ_AGE_1 > 840 ~ NA_integer_,
    PROJ_AGE_1 <= 840 & PROJ_AGE_1 > 250 ~ as.integer(250),
    TRUE ~ PROJ_AGE_1),
    AGE.ROUNDED = floor(PROJ_AGE_1 / 10) * 10)

age.df <- vri %>%
  dplyr::mutate(AREA.METRES = sf::st_area(.)) %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(AGE.ROUNDED) %>%
  dplyr::summarize(AREA.BY.GRP = as.numeric(sum(AREA.METRES, na.rm = TRUE))/10000000) %>%
  dplyr::filter(!is.na(AGE.ROUNDED))


max.area = max(age.df$AREA.BY.GRP, na.rm = TRUE)

ggplot(age.df, aes(x = AGE.ROUNDED, y = AREA.BY.GRP)) +
  geom_bar(stat = "identity", fill = "goldenrod") +
  labs(
    # title = "Age Distribution of Forest Stands by Area",
    x = "Age (years)",
    y = "Total Area (1000s of hectares)",
  ) +
  scale_x_continuous(breaks = seq(0, 250, by = 10)) +
  scale_y_continuous(breaks = seq(0, max.area, by = 10)) +
  theme(
    # text = element_text(family = "anton"),
    panel.background = element_rect("transparent"),
    panel.grid.major = element_line(colour = "grey88"),
    # panel.grid.minor = element_line(colour = "grey88"),
    # axis.ticks = element_blank(),
    axis.text = element_text(size = 14),
    # axis.ticks = element_line(size = 0.5),
    axis.title.x = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    plot.title = element_text(size = 45, hjust = 0.5),
  )

fig.name <- "C:/Users/clarmour/Documents/GitHub/PEMtestr/summary_report/figures/age_by_area.png"

```

```{r}

#| fig-cap: "Figure 4: A map of forest age across [AOI_name]. Polygons with an age of 250 years or greater were clamped to 250 years for visualization purposes."
#| fig-width: 6
#| fig-height: 4

ggplot() +
  tidyterra::geom_spatvector(data = dplyr::filter(vri, PROJ_AGE_1 > 0), mapping = aes(fill = AGE.ROUNDED), color = "transparent") +
  viridis::scale_fill_viridis(direction = -1, option = "D") +
  labs(fill = "Forest Age (years)") +
    ggspatial::annotation_north_arrow(
    location = "tr", # Position: bottom-right
    which_north = "true", # True North
    pad_x = unit(0.5, "cm"), # Padding from x-axis
    pad_y = unit(0.5, "cm"), # Padding from y-axis
    style = ggspatial::north_arrow_fancy_orienteering() # Style of the arrow
  )


```

```{r species-dist, echo = FALSE}

### SPECIES DISTRIBUTION

species.df <- vri %>%
  dplyr::filter(!is.na(SPECIES_CD_1)) %>%
  dplyr::mutate(AREA.METRES = sf::st_area(.)) %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(SPECIES_CD_1) %>%
  dplyr::summarize(AREA.BY.GRP = as.numeric(sum(AREA.METRES, na.rm = TRUE))/10000000) %>%
  dplyr::mutate(SPECIES_CD_1 = as.factor(SPECIES_CD_1)) %>%
  dplyr::arrange(desc(AREA.BY.GRP)) %>%
  dplyr::filter(AREA.BY.GRP >= 1)

max.area = max(species.df$AREA.BY.GRP, na.rm = TRUE)

ggplot(species.df, aes(x = reorder(SPECIES_CD_1, -AREA.BY.GRP), y = AREA.BY.GRP)) +
  geom_bar(stat = "identity", fill = pal(8)[4]) +
  labs(
    # title = "Age Distribution of Forest Stands by Area",
    x = "Species Code",
    y = "Total Area (1000s of hectares)",
  ) +
  # scale_x_continuous(breaks = seq(0, 250, by = 10)) +
  scale_y_continuous(breaks = seq(0, max.area, by = floor(max.area/10))) +
  theme(
    # text = element_text(family = "anton"),
    panel.background = element_rect("transparent"),
    panel.grid.major = element_line(colour = "grey88"),
    # panel.grid.minor = element_line(colour = "grey88"),
    # axis.ticks = element_blank(),
    # axis.text = element_text(size = 24),
    # axis.ticks = element_line(size = 0.5),
    # axis.title.x = element_text(size = 32, face = "bold"),
    # axis.title.y = element_text(size = 32, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    plot.title = element_text(size = 54, hjust = 0.5),
  )
```
