---
title: "PrepR Summary Report"
format: PrettyPDF-pdf
fig.height: 4.5
fig.width: 6.5
editor_options: 
  chunk_output_type: console
---
```{r setup, include = FALSE}
require(tidyverse)
require(magrittr)
require(terra)
require(sf)
require(pals)
require(ochRe)
require(showtext)
require(ggtext)
require(gghighlight)

```
```{r graphics-setup, include = FALSE}

pal <- colorRampPalette(ochRe::ochre_palettes[["namatjira_qual"]])
font_add_google('Anton', 'anton')
showtext_auto()

```
This report provides several metrics of interest over the geographic extent of the AOI, including the distribution of forest age, dominant species, common disturbances, topographic variables, and others. It also summarizes key outputs from the prepR workflow and briefly describes their purpose in the modelling process. More details on the data, methodology, and results from the prepR workflow can be found in the PEM Manual.

```{r aoi-inset, echo = FALSE}

aoi <- st_read("C:/Users/ccarmour.stu/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/aoi.gpkg", quiet = TRUE)

bc_outline <- st_read("C:/Users/ccarmour.stu/Documents/GitHub/PEMtestr/summary_report/bc_outline.gpkg", quiet = TRUE)

ggplot() +
  geom_sf(data = st_geometry(bc_outline)) +
  geom_sf(data = st_geometry(aoi), color = "red")

```

```{r bec-map, echo = FALSE}

bec <- st_read("C:/Users/ccarmour.stu/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/bec.gpkg", quiet = TRUE) %>%
  dplyr::mutate(BGC_LABEL = str_replace_all(BGC_LABEL, " ", ""))

ggplot(data = bec, mapping = aes(fill = BGC_LABEL))+
  geom_sf() +
  scale_fill_manual(values = pal(6)) +
  labs(fill = "BEC Subzone/Variant")

```


```{r forest-age, echo = FALSE}

vri <- st_read("C:/Users/ccarmour.stu/OneDrive - Government of BC/PEM/AOIs/DateCreek_AOI/0_raw_inputs/base_layers/vri_detail.gpkg", quiet = TRUE) %>%
  dplyr::mutate(BEC_LABEL = str_c(BEC_ZONE_CODE, BEC_SUBZONE),
                BEC_LABEL = case_when(
                  !is.na(BEC_VARIANT) ~ str_c(BEC_LABEL, BEC_VARIANT),
                  TRUE ~ BEC_LABEL))

### FOREST AGE

## Extract forest age from PROJ_AGE_1 variable
## Calculate area in hectares, drop geometry
## Clamp between 0-250 years and bin by decade
## Group by age class and sum total area in hectares
age.df <- vri %>%
  dplyr::mutate(AREA.METRES = sf::st_area(.)) %>%
  sf::st_drop_geometry() %>%
  dplyr::mutate(PROJ_AGE_1 = case_when(
    PROJ_AGE_1 == 0 ~ NA_integer_,
    PROJ_AGE_1 > 840 ~ NA_integer_,
    PROJ_AGE_1 <= 840 & PROJ_AGE_1 > 250 ~ as.integer(250),
    TRUE ~ PROJ_AGE_1),
    AGE.ROUNDED = floor(PROJ_AGE_1 / 10) * 10) %>%
  dplyr::group_by(AGE.ROUNDED) %>%
  dplyr::summarize(AREA.BY.GRP = as.numeric(sum(AREA.METRES, na.rm = TRUE))/10000000) %>%
  dplyr::filter(!is.na(AGE.ROUNDED))

max.area = max(age.df$AREA.BY.GRP, na.rm = TRUE)

## plotting in R
ggplot(age.df, aes(x = AGE.ROUNDED, y = AREA.BY.GRP)) +
  geom_bar(stat = "identity", fill = pal(8)[6]) +
  labs(
    # title = "Age Distribution of Forest Stands by Area",
    x = "Age (years)",
    y = "Total Area (1000s of hectares)",
  ) +
  scale_x_continuous(breaks = seq(0, 250, by = 10)) +
  scale_y_continuous(breaks = seq(0, max.area, by = 10)) +
  theme(
    # text = element_text(family = "anton"),
    panel.background = element_rect("transparent"),
    panel.grid.major = element_line(colour = "grey88"),
    # panel.grid.minor = element_line(colour = "grey88"),
    # axis.ticks = element_blank(),
    # axis.text = element_text(size = 24),
    # axis.ticks = element_line(size = 0.5),
    # axis.title.x = element_text(size = 32, face = "bold"),
    # axis.title.y = element_text(size = 32, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    plot.title = element_text(size = 54, hjust = 0.5),
  )

```

```{r species-dist, echo = FALSE}

### SPECIES DISTRIBUTION

species.df <- vri %>%
  dplyr::filter(!is.na(SPECIES_CD_1)) %>%
  dplyr::mutate(AREA.METRES = sf::st_area(.)) %>%
  sf::st_drop_geometry() %>%
  dplyr::group_by(SPECIES_CD_1) %>%
  dplyr::summarize(AREA.BY.GRP = as.numeric(sum(AREA.METRES, na.rm = TRUE))/10000000) %>%
  dplyr::mutate(SPECIES_CD_1 = as.factor(SPECIES_CD_1)) %>%
  dplyr::arrange(desc(AREA.BY.GRP)) %>%
  dplyr::filter(AREA.BY.GRP >= 1)

max.area = max(species.df$AREA.BY.GRP, na.rm = TRUE)

ggplot(species.df, aes(x = reorder(SPECIES_CD_1, -AREA.BY.GRP), y = AREA.BY.GRP)) +
  geom_bar(stat = "identity", fill = pal(8)[4]) +
  labs(
    # title = "Age Distribution of Forest Stands by Area",
    x = "Species Code",
    y = "Total Area (1000s of hectares)",
  ) +
  # scale_x_continuous(breaks = seq(0, 250, by = 10)) +
  scale_y_continuous(breaks = seq(0, max.area, by = floor(max.area/10))) +
  theme(
    # text = element_text(family = "anton"),
    panel.background = element_rect("transparent"),
    panel.grid.major = element_line(colour = "grey88"),
    # panel.grid.minor = element_line(colour = "grey88"),
    # axis.ticks = element_blank(),
    # axis.text = element_text(size = 24),
    # axis.ticks = element_line(size = 0.5),
    # axis.title.x = element_text(size = 32, face = "bold"),
    # axis.title.y = element_text(size = 32, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 0.5),
    plot.title = element_text(size = 54, hjust = 0.5),
  )
```

